---
title: "Covid91 vaccine study Final2023"
author: "Tito Lopez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

$H_0$ The covid-921 vaccine does not exhibit a significant difference in effectiveness between males

$H_A$ The covid-921 vaccine does exhibit a significant difference in effectiveness between males

```{r}
library(dplyr)
males = filter(FinalData, FinalData$sex=="M")
```


## Methods

Males is categorical and covid-921 is explanatory, therefore we use cat~cat.

## Graphical Results

```{r}
barchartGC(~infected + treatment,data=males)
barchartGC(~infected + treatment,data=males, type="percent")
```


## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=males)
rowPerc(table1)
colPerc(table1)
```


## Inferental Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

## Females

$H_0$ The covid-921 vaccine does not exhibit a significant difference in effectiveness between female

$H_A$ The covid-921 vaccine does exhibit a significant difference in effectiveness between female

```{r}
library(dplyr)
female = filter(FinalData, FinalData$sex=="F")
```

## Methods



## Graphical Results



## Numerical Results



## Inferental Results



## LGBTQ

$H_0$ The covid-921 vaccine does not exhibit a significant difference in effectiveness between lgbtq

$H_A$ The covid-921 vaccine does exhibit a significant difference in effectiveness between lgbtq

```{r}
library(dplyr)
lgbtq = filter(FinalData, FinalData$LGBTQ=="gay")
```

## Methods



## Graphical Results



## Numerical Results



## Inferental Results



## Druggies

$H_0$ The covid-921 vaccine does not exhibit a significant difference in effectiveness between druggies

$H_A$ The covid-921 vaccine does exhibit a significant difference in effectiveness between druggies

```{r}
library(dplyr)
druggies = filter(FinalData, FinalData$DrugUser=="Yes")
```

## Methods



## Graphical Results



## Numerical Results



## Inferental Results



# Overall Results and Conclusions

